---
title: "Explore_Water_Supply_Database"
author: "KEC"
date: "2025-01-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("setup.R")

```


First, import some basic NPS information.

```{r}

park_of_interest <- "CAHA"

all_parks <- getParkSystem()

nc_parks <- all_parks %>% dplyr::filter(STATE == "NC")
nc_counties <- tigris::counties(state = "NC")

park_counties <- nc_counties %>%
  st_filter(nc_parks)

mapview(park_counties) +
  mapview(nc_parks, col.regions = "red") +
#  mapview(nc_wells, col.regions = "black")
  mapview(buildings %>% st_centroid(),
          zcol = "BLDGNAME")
  
# Select specific parks for analysis using filter
select_parks <- all_parks %>%
  dplyr::filter(UNIT_CODE == park_of_interest)
                STATE == "NC")

# Get buildings within 5000m (5km) buffer of select parks
buildings <- getBuildings(park_boundary = select_parks %>%
                            st_buffer(5000))
```

Now, plot select park and buildings. 

```{r}

a <-
  mapview(select_parks,
        alpha.regions = 0.2,
        col.regions = "seagreen",
        layer.name = "Park Boundary") +
  mapview(buildings %>% st_centroid(),
        col.regions = "tomato",
        layer.name = "Buildings")
a

```


Now import the Water Supply Database

You will likely need to modify the table path with the location of the database on your system -- unless a symlink has been created in the project folder. To add a symlink, open the terminal tab below and type:

ln -s “directory_you_want_to_link” “directory_where_link_will_live”

KEC: For me, this would be:
ln -s "/Users/kcognac/Library/CloudStorage/OneDrive-SharedLibraries-Colostate/Willi,Katie - nps_water_vulnerability/data" "/Users/kcognac/Desktop/Repositories/NPS_WSD"


```{r}
table_path <- "data/Water_Supply_Systems/NPS_Water_Systems_Database_Joined.xlsx"

# Supply table 
supply_table <- read_excel(table_path, sheet = 1, skip = 1, na = "NA") %>%
  janitor::clean_names() %>%
  dplyr::filter(park_unit %in% nc_parks$UNIT_CODE)

# Create copy of source table as sf_object (i.e., with geospatial attributes). 
# During this process, it is required that rows without location data are
# dropped. Circle back to this later.
source_table_locs <- supply_table %>%
  drop_na(c("source_longitude", "source_longitude")) %>%
  st_as_sf(., 
           coords = (c("source_longitude","source_latitude")), 
           crs = 4326,
           remove = FALSE) 



select_sources <- source_table_locs %>%
  dplyr::filter(park_unit %in% select_parks$UNIT_CODE)

b <- a + mapview(select_sources,
                 col.regions = "orange",
                 layer.name = "Water Sources")

b
        
```


Depending on the state, import well data
```{r}

nevada_wells <- read_csv("/Users/kcognac/Desktop/Well_Driller_Reports.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326,
           remove = FALSE)

wells_near_parks <- nevada_wells %>%
  st_intersection(select_parks %>%
                    st_buffer(5000) %>%
                    st_transform(crs = st_crs(nevada_wells)))

c <- b + mapview(wells_near_parks,
                 col.regions = "dodgerblue",
                 cex = 3)

c

```


There is an existing harmonized database for water rights (HarDWR) that can be useful for quickly finding water rights. Note, this was made about a year ago (and counting), so it is not guaranteed to be inclusive. But, it's a good first pass. I've included an example of Wyoming below, but see if you can figure out how to get the Nevada one loaded. Note, you might need to navigate to the folder on onedrive and unzip it manually first.

```{r HarDWR}


hardwr_wyoming <- read_csv("data/all/HarDWR/stateWaterRightsRaw/WyomingRights/Wyoming_HarDWR_Raw.csv") %>%
  drop_na(c("lon","lat")) %>%
  st_as_sf(., 
           coords = (c("lon","lat")), 
           crs = 4326,
           remove = FALSE)


```


There is also a harmonized database of wells across the United States. Again, this is only as up to date as when the database was created. However, it might be useful for states where it is a challenge to find wells.  I've included some examples, but again, navigate to the directory using onedrive and explore if needed.

```{r USGWD}

wyoming_wells <- read_csv("data/all/Water_Supply/USGWD-Tabular/USGWD_Wyoming.csv") %>%
  drop_na(c("Longitude","Latitude")) %>%
  st_as_sf(., 
           coords = (c("Longitude","Latitude")), 
           crs = 4326,
           remove = FALSE)


texas_wells <- read_csv("data/all/Water_Supply/USGWD-Tabular/USGWD_Texas.csv")%>%
  drop_na(c("Longitude","Latitude")) %>%
  st_as_sf(., 
           coords = (c("Longitude","Latitude")), 
           crs = 4326,
           remove = FALSE)

newmexico_wells <- read_csv("data/all/Water_Supply/USGWD-Tabular/USGWD_New_Mexico.csv")%>%
  drop_na(c("Longitude","Latitude")) %>%
  st_as_sf(., 
           coords = (c("Longitude","Latitude")), 
           crs = 4326,
           remove = FALSE)


colorado_wells <- read_csv("data/all/Water_Supply/USGWD-Tabular/USGWD_Colorado.csv")%>%
  drop_na(c("Longitude","Latitude")) %>%
  st_as_sf(., 
           coords = (c("Longitude","Latitude")), 
           crs = 4326,
           remove = FALSE)

utah_wells <- read_csv("data/all/Water_Supply/USGWD-Tabular/USGWD_Utah.csv")%>%
  drop_na(c("Longitude","Latitude")) %>%
  st_as_sf(., 
           coords = (c("Longitude","Latitude")), 
           crs = 4326,
           remove = FALSE)


# NORTH CAROLINA

nc_wells <- read_csv("data/all/Water_Supply/USGWD-Tabular/USGWD_North_Carolina.csv")%>%
  drop_na(c("Longitude","Latitude")) %>%
  st_as_sf(., 
           coords = (c("Longitude","Latitude")), 
           crs = 4326,
           remove = FALSE)

# From irma datastore: https://irma.nps.gov/DataStore/Reference/Profile/1016845
CAHA_wqgis_path <- "data/Water_Supply_Systems/data_sources/North_Carolina/caha_wqgis"
CAHA_layers <- st_layers(CAHA_wqgis_path) # see all layers
CAHA_dg <- st_read(path, layer = "idg_station", crs = 32618)
mapview(CAHA_dg)

```

  